### 2024-05-26 getting-started.qmd
20240526 - subset

>>>>

## Subset {#sec-subset}

To subset a data frame, use brackets to specify the subset of rows and columns to keep, where the value/vector before the comma specifies the rows to keep, and the value/vector after the comma specifies the columns to keep:

```{r, eval = FALSE}
dataframe[rowsToKeep, columnsToKeep]
```

You can subset by using any of the following:

- numeric indices of the rows/columns to keep (or drop)
- names of the rows/columns to keep (or drop)
- values of `TRUE` and `FALSE` corresponding to which rows/columns to keep

### One Variable {#sec-subsetOneVariable}

To subset one variable, use the following syntax:

```{r}
nfl_players$display_name
```

or:

```{r}
nfl_players[,"display_name"]
```

### Particular Rows of One Variable {#sec-subsetParticularRowsOneVariable}

To subset one variable, use the following syntax:

```{r}
nfl_players$display_name[which(nfl_players$position_group == "RB")]
```

or:

```{r}
nfl_players[which(nfl_players$position_group == "RB"), "display_name"]
```

### Particular Columns (Variables) {#sec-subsetParticularColumns}

To subset particular columns/variables, use the following syntax:

#### Base `R` {#sec-subsetParticularColumnsBaseR}

```{r}
subsetVars <- c("status","display_name","college_name")

nfl_players[,c(1,2,9)]
nfl_players[,c("status","display_name","college_name")]
nfl_players[,subsetVars]
```

Or, to drop columns:

```{r}
dropVars <- c("status","college_name")

nfl_players[,-c(1,9)]
nfl_players[,!(names(nfl_players) %in% c("status","college_name"))]
nfl_players[,!(names(nfl_players) %in% dropVars)]
```

#### Tidyverse {#sec-subsetParticularColumnsTidyverse}

```{r}
nfl_players %>%
  select(status, display_name, college_name)

nfl_players %>%
  select(status:college_name)

nfl_players %>%
  select(all_of(subsetVars))
```

Or, to drop columns:

```{r}
nfl_players %>%
  select(-status, -college_name)

nfl_players %>%
  select(-c(status:college_name))

nfl_players %>%
  select(-all_of(dropVars))
```

### Particular Rows {#sec-subsetParticularRows}

To subset particular rows, use the following syntax:

#### Base `R` {#sec-subsetParticularRowsBaseR}

```{r}
subsetRows <- c(1,3,5)

nfl_players[c(1,3,5),]
nfl_players[subsetRows,]
nfl_players[which(nfl_players$position_group == "RB"),]
```

#### Tidyverse {#sec-subsetParticularRowsTidyverse}

```{r}
nfl_players %>%
  filter(position_group == "RB")

nfl_players %>%
  filter(position_group == "RB", weight <= 250)

nfl_players %>%
  filter(position_group == "RB" | weight >= 250)
```

### Particular Rows and Columns {#sec-subsetParticularRowsAndColumns}

To subset particular rows and columns, use the following syntax:

#### Base `R` {#sec-subsetParticularRowsAndColumnsBaseR}

```{r}
nfl_players[c(1,3,5), c(1,2,3)]
nfl_players[subsetRows, subsetVars]
nfl_players[which(nfl_players$position_group == "RB"), subsetVars]
```

#### Tidyverse {#sec-subsetParticularRowsAndColumnsTidyverse}

```{r}
nfl_players %>%
  filter(position_group == "RB") %>%
  select(all_of(subsetVars))
```
